---
title: "Data cleaning"
date: "2023-07-31"
output: html_document
---
#library
```{r}
library(readr)
library(geosphere)
library(dplyr)
library(ggplot2)
library(gridExtra)
library(grid)
library(lme4)
library(lmerTest)
library(ggspatial)
library(cowplot)
library(rnaturalearth)
library(ggmap)
```

#data
```{r}
rm(list=ls())

#If you are using the input data from Github "Wind-drift-compensation-analysis/Input/GPS/HMM state classified/"
csv_files <- list.files(path = "~/Documents/yodalab/paper/JAB/data/State_Classified/", 
                        pattern = "\\.csv$", full.names = TRUE)
for (file in csv_files) {
  var_name <- tools::file_path_sans_ext(basename(file))
  assign(var_name, read_csv(file), envir = .GlobalEnv)
}

#If you are using the single file combined all years
# data <- read_csv(
#   "~/Documents/yodalab/paper/JAB/data/State_Classified/HMM_classified.csv",
#   show_col_types = FALSE
# )
# 
# # Make GPS2013-GPS2022 from the year column
# for (yr in 2013:2022) {
#   
#   assign(
#     paste0("GPS", yr),
#     data %>% filter(year == yr),
#     envir = .GlobalEnv
#   )
# }
```

#Cleaning1: make the state of the last row to "Homing"
Now the last row is NA, so need to fix it to homing
```{r}
update_last_row_conditionally <- function(df) {
  df %>%
    group_by(ID) %>%
    mutate(State = if_else(row_number() == n() & coast_distance >= 5000, "Fast_Homing", State)) %>%
    ungroup()
}
GPS2013 <- update_last_row_conditionally(GPS2013)
GPS2014 <- update_last_row_conditionally(GPS2014)
GPS2015 <- update_last_row_conditionally(GPS2015)
GPS2016 <- update_last_row_conditionally(GPS2016)
GPS2017 <- update_last_row_conditionally(GPS2017)
GPS2018 <- update_last_row_conditionally(GPS2018)
GPS2019 <- update_last_row_conditionally(GPS2019)
GPS2020 <- update_last_row_conditionally(GPS2020)
GPS2021 <- update_last_row_conditionally(GPS2021)
GPS2022 <- update_last_row_conditionally(GPS2022)
```

#Integrate all data to 1 dataframe
```{r}
all_data <- rbind(GPS2013, GPS2014, GPS2015, GPS2016, GPS2017, GPS2018, GPS2019, GPS2020, GPS2021, GPS2022)

write_csv(all_data, "~/Documents/yodalab/paper/JAB/data/not_for_upload/HMM_added/output/all_state_data.csv")
```


#Cleaning1' adopt original_state for selected trips
In the actual data, HMM classification was performed both with and without coastal exclusion. The results with coastal exclusion were stored in the "State" column, and the results without coastal exclusion were stored in the "original_state" column.

To improve accuracy, each trip was visually inspected. For trips where errors were observed (e.g., Searching segments mixed into Homing), and where the original_state classification (i.e., without coastal exclusion) was found to be more accurate, that classification was manually adopted instead.
```{r}
# List of Trip IDs to replace State with original_state
# "2013_9A67016_12_trip","2014_9A79458_2_trip","2014_K14B89P_5_trip",
# "2022_9B19815_43_trip","2022_9B14746_47_trip","2022_9B14746_48_trip",
# "2019_9A58931_15_trip","2013_K13C40P_2_trip","2014_9A66707_11_trip", "2014_K14B89P_7_trip"
ids_to_replace <- c("2013_9A67016_12_trip","2014_9A79458_2_trip","2014_K14B89P_5_trip","2022_9B19815_43_trip","2022_9B14746_47_trip","2022_9B14746_48_trip","2019_9A58931_15_trip","2013_K13C40P_2_trip","2014_9A66707_11_trip","2014_K14B89P_7_trip")
# Replace State with original_state for the listed IDs
mask <- all_data$ID %in% ids_to_replace & !is.na(all_data$State)
all_data$State[mask] <- all_data$original_state[mask]

# List of Trip IDs to remove
# "2022_9B00478_76_trip", "2013_K13C40P_5_trip"
# These trips are excluded because their departure points are
# located too far toward the upper or lower boundary of the study area.
ids_to_remove <- c("2022_9B00478_76_trip", "2013_K13C40P_5_trip")
# Remove rows belonging to the listed IDs
all_data <- all_data[!all_data$ID %in% ids_to_remove, ]
```

#Cleaning2: Extract the homing data on sea, considering non-flight state
When a "Slow" segment is interspersed within Homing, the bird is not flying, which causes inaccurate speed calculations.
To address this, Homing_ID column is added to split Homing into separate segments when a Slow segment is interspersed, then only Homing data is extracted.
```{r}
# 1. For each ID, identify the first occurrence of Fast_Homing

all_data <- all_data %>%
  filter(!is.na(State))
filtered_data <- all_data %>%
  group_by(ID) %>%
  filter(cumsum(State == "Fast_Homing") > 0)
length(unique(filtered_data$ID))
#[1] 1368

# 2. Create a new HomingID column and assign a unique ID to each Homing segment

homing_id <- rle(as.character(filtered_data$State))$lengths
homing_id <- rep(seq_along(homing_id), homing_id)

filtered_data <- filtered_data %>%
  mutate(new_HomingID = ifelse(State == "Fast_Homing", 
                               cumsum(State == "Fast_Homing" & lag(State, default="") != "Fast_Homing"), 
                               NA_integer_)) %>%
  mutate(HomingID = ifelse(!is.na(new_HomingID), 
                           paste0(ID, "_", new_HomingID), 
                           NA_character_)) %>%
  dplyr::select(-new_HomingID)

check <- filtered_data %>% 
  dplyr::select(ID, State, HomingID)

length(unique(filtered_data$HomingID))
#[1] 1787

# Remove unnecessary columns and rows where HomingID is NA
filtered_data <- filtered_data %>%
  filter(!is.na(HomingID))
# Remove HomingIDs that contain only a single row
filtered_data <- filtered_data %>%
  group_by(HomingID) %>%
  filter(!is.na(HomingID) & n() > 1) %>%
  ungroup()

all_homing_data <- filtered_data

length(unique(all_homing_data$ID))
#[1] 1296
length(unique(filtered_data$HomingID))
#[1] 1648
```

#Cleaning3:Remove the ID less than 5 points
```{r}
removed_IDs <- all_homing_data %>%
  group_by(ID) %>%
  filter(State == "Fast_Homing") %>% 
  filter(n() <= 5) %>%
  distinct(ID) %>%
  pull(ID)

cat("Removed IDs:", paste(removed_IDs, collapse = ", "), "\n")

all_homing_data <- all_homing_data %>%
  filter(!(ID %in% removed_IDs))
length(unique(all_homing_data$ID))
```
#Output and read again
```{r}
write_csv(all_homing_data, "~/Documents/yodalab/paper/JAB/data/not_for_upload/HMM_added/output/all_sea_data.csv")

rm(list=ls())
all_sea_data <- read_csv("~/Documents/yodalab/paper/JAB/data/not_for_upload/HMM_added/output/all_sea_data.csv")
```

#Visualize all 
```{r}
# Plot all GPS tracks on a map
japan_map <- map_data("world", region = "Japan")
location_nest_data <- data.frame(lon = 141.5576031, lat = 40.5386062)
x_range <- c(141.1705, 144.0207)
y_range <- c(38.82596, 44.48974)

ggplot() +
  geom_polygon(data = japan_map, aes(x = long, y = lat, group = group), fill = "white", color = "black") +
  geom_path(data = all_sea_data, aes(x = lon, y = lat, group = ID), size = 0.5, alpha = 0.3, colour = "steelblue") +
  geom_point(data = all_sea_data, aes(x = lon, y = lat, colour = State), size = 0.5, alpha = 0.5) +
  # Add nest location
  geom_point(data = location_nest_data, aes(x = lon, y = lat), color = "black", size = 3) +
  # Use coord_map() instead of coord_cartesian() to correct map distortion
  coord_map(xlim = x_range, ylim = y_range) +
  theme_classic()
```

#Cleaning4: remove the wierld trip manually
```{r}
#check <- all_sea_data %>% filter(ID %in% c("2015_9A32131_5_trip", "2014_K14B89P_7_trip"))
check <- all_sea_data %>% filter(ID %in% c("2022_9B07922_12_trip"))


  japan_map <- map_data("world", region = "Japan")
  location_nest_data <- data.frame(lon = 141.5576031, lat = 40.5386062)
  x_range <- c(141.1705, 146.0207)
  y_range <- c(38.82596, 43.48974)

  ggplot() +
        geom_polygon(data = japan_map, aes(x = long, y = lat, group = group), fill = "white", color = "black") +
        geom_point(data = check, aes(x = lon, y = lat, colour = original_state), size = 1) +
        geom_path(size = 1.5, alpha = 0.5, colour = "snow2") +
        geom_point(data = location_nest_data, aes(x = lon, y = lat), color = "black", size = 2) +
        coord_cartesian(xlim = x_range, ylim = y_range) +
        theme_classic()

################################################################################
## Remove trips whose beeline from the departure point to the nest passes over land
# "2013_9A67016_4_trip", "2022_9B00478_11_trip", "2022_9B04850_1_trip"
 ################################################################################
# List of removing IDs
exclude_ids <- c("2013_9A67016_4_trip", "2022_9B00478_11_trip", "2022_9B04850_1_trip")
all_sea_data <- all_sea_data %>%
  filter(!(ID %in% exclude_ids))

################################################################################
## Remove the last point of trips where the bird is still flying at the final GPS point
# "2022_9B07922_12_trip"
################################################################################
all_sea_data <- all_sea_data %>%
  group_by(ID) %>%
  mutate(row_number = row_number()) %>%
  filter(!(ID == "2022_9B07922_12_trip" & row_number == max(row_number))) %>%
  dplyr::select(-row_number)

######################################################################
#Remove the exceptionaly long trip
#2015_9A32131_5_trip
######################################################################

exclude_ids <- c("2015_9A32131_5_trip")

all_sea_data <- all_sea_data %>%
  filter(!(ID %in% exclude_ids))
```

#Visualize all 
```{r}
# Plot all GPS tracks on a map
japan_map <- map_data("world", region = "Japan")
location_nest_data <- data.frame(lon = 141.5576031, lat = 40.5386062)
x_range <- c(141.1705, 144.0207)
y_range <- c(38.82596, 44.48974)

ggplot() +
  geom_polygon(data = japan_map, aes(x = long, y = lat, group = group), fill = "white", color = "black") +
  geom_path(data = all_sea_data, aes(x = lon, y = lat, group = ID), size = 0.5, alpha = 0.3, colour = "steelblue") +
  geom_point(data = all_sea_data, aes(x = lon, y = lat, colour = State), size = 0.5, alpha = 0.5) +
  # Add nest location
  geom_point(data = location_nest_data, aes(x = lon, y = lat), color = "black", size = 3) +
  # Use coord_map() instead of coord_cartesian() to correct map distortion
  coord_map(xlim = x_range, ylim = y_range) +
  theme_classic()
```

#Cleaning5: Exclude the data missing age
```{r}
# Remove rows where age is NA
all_sea_data <- all_sea_data %>%
  filter(!is.na(age))
```

#Visualize all 
```{r}
# Plot all GPS tracks on a map
japan_map <- map_data("world", region = "Japan")
location_nest_data <- data.frame(lon = 141.5576031, lat = 40.5386062)
x_range <- c(141.1705, 144.0207)
y_range <- c(38.82596, 44.48974)

ggplot() +
  geom_polygon(data = japan_map, aes(x = long, y = lat, group = group), fill = "white", color = "black") +
  geom_path(data = all_sea_data, aes(x = lon, y = lat, group = ID), size = 0.5, alpha = 0.3, colour = "steelblue") +
  geom_point(data = all_sea_data, aes(x = lon, y = lat, colour = State), size = 0.5, alpha = 0.5) +
  # Add nest location
  geom_point(data = location_nest_data, aes(x = lon, y = lat), color = "black", size = 3) +
  # Use coord_map() instead of coord_cartesian() to correct map distortion
  coord_map(xlim = x_range, ylim = y_range) +
  theme_classic()
```

#Output
```{r}
write_csv(all_sea_data, "~/Documents/yodalab/paper/JAB/data/not_for_upload/HMM_added/output/all_sea_data.csv")
```

