---
title: "HMM2"
output: html_document
date: "2023-09-12"
---
HMM classification cannot classify all IDs in a single run.
It is necessary to run the classification multiple times and verify
that all IDs have been successfully assigned a state.
Run this script as a continuation of New5_HMM1.Rmd.

#14. read GPS data
data before running HMM
```{r}
csv_files <- list.files(path = "~/yodalab/BTG_Trj_U/analyze_data/middle_data/Added/No5_coast_added/after/", 
                        pattern = "\\.csv$", full.names = TRUE)
for (file in csv_files) {
  var_name <- tools::file_path_sans_ext(basename(file))
  assign(var_name, read_csv(file), envir = .GlobalEnv)
}
```

#15. read HMM classified data
```{r}
HMM2013 <- read_csv("~/Documents/yodalab/data/middle_data/Added/No5_HMM_added/HMM_GPS2013.csv")
HMM2014 <- read_csv("~/Documents/yodalab/data/middle_data/Added/No5_HMM_added/HMM_GPS2014.csv")
HMM2015 <- read_csv("~/Documents/yodalab/data/middle_data/Added/No5_HMM_added/HMM_GPS2015.csv")
HMM2016 <- read_csv("~/Documents/yodalab/data/middle_data/Added/No5_HMM_added/HMM_GPS2016.csv")
HMM2017 <- read_csv("~/Documents/yodalab/data/middle_data/Added/No5_HMM_added/HMM_GPS2017.csv")
HMM2018 <- read_csv("~/Documents/yodalab/data/middle_data/Added/No5_HMM_added/HMM_GPS2018.csv")
HMM2019_1 <- read_csv("~/Documents/yodalab/data/middle_data/Added/No5_HMM_added/HMM_GPS2019_1.csv")
HMM2019_2 <- read_csv("~/Documents/yodalab/data/middle_data/Added/No5_HMM_added/HMM_GPS2019_2.csv")
HMM2019 <- rbind(HMM2019_1, HMM2019_2)
HMM2020_1 <- read_csv("~/Documents/yodalab/data/middle_data/Added/No5_HMM_added/HMM_GPS2020_1.csv")
HMM2020_2 <- read_csv("~/Documents/yodalab/data/middle_data/Added/No5_HMM_added/HMM_GPS2020_2.csv")
HMM2020 <- rbind(HMM2020_1, HMM2020_2)
HMM2021_1 <- read_csv("~/Documents/yodalab/data/middle_data/Added/No5_HMM_added/HMM_GPS2021_1.csv")
HMM2021_2 <- read_csv("~/Documents/yodalab/data/middle_data/Added/No5_HMM_added/HMM_GPS2021_2.csv")
HMM2021 <- rbind(HMM2021_1, HMM2021_2)
HMM2022_1 <- read_csv("~/Documents/yodalab/data/middle_data/Added/No5_HMM_added/HMM_GPS2022_1.csv")
HMM2022_2 <- read_csv("~/Documents/yodalab/data/middle_data/Added/No5_HMM_added/HMM_GPS2022_2.csv")
HMM2022_3 <- read_csv("~/Documents/yodalab/data/middle_data/Added/No5_HMM_added/HMM_GPS2022_3.csv")
HMM2022_4 <- read_csv("~/Documents/yodalab/data/middle_data/Added/No5_HMM_added/HMM_GPS2022_4.csv")
HMM2022_5 <- read_csv("~/Documents/yodalab/data/middle_data/Added/No5_HMM_added/HMM_GPS2022_5.csv")
HMM2022_6 <- read_csv("~/Documents/yodalab/data/middle_data/Added/No5_HMM_added/HMM_GPS2022_6.csv")
HMM2022 <- rbind(HMM2022_1,HMM2022_2,HMM2022_3,HMM2022_4,HMM2022_5,HMM2022_6)
```

#16. Match states to GPS data (1st round)
After this step, the State column for each ID will contain:
  - Success: one of the 4 states (Fast_Outbound, Fast_Middle,
             Fast_Homing, Slow), with NA only at the last point
  - Failure: NA for all points
```{r}
update_state <- function(year) {
  year_string <- as.character(year)
  data_name <- paste0(input_name, year_string)
  model_name <- paste0(state_name, year_string)
  
  data <- get(data_name)
  model <- get(model_name)
  
  data$State <- NA
  for (i in 1:length(model$ID)) {
    index <- which(data$ID == model$ID[i] & data$time == model$time[i])
    data$State[index] <- model$State[i]
  }
  
  assign(data_name, data, envir = .GlobalEnv)
  
}

input_name <- "GPS"
state_name <- "HMM"
years <- 2013:2022
lapply(years, update_state)
```

#17. Output
In case your laptop will die
```{r}
write_csv(GPS2013,"~/yodalab/BTG_Trj_U/analyze_data/middle_data/Added/No5_HMM_added/temporary/GPS2013.csv")
write_csv(GPS2014,"~/yodalab/BTG_Trj_U/analyze_data/middle_data/Added/No5_HMM_added/temporary/GPS2014.csv")
write_csv(GPS2015,"~/yodalab/BTG_Trj_U/analyze_data/middle_data/Added/No5_HMM_added/temporary/GPS2015.csv")
write_csv(GPS2016,"~/yodalab/BTG_Trj_U/analyze_data/middle_data/Added/No5_HMM_added/temporary/GPS2016.csv")
write_csv(GPS2017,"~/yodalab/BTG_Trj_U/analyze_data/middle_data/Added/No5_HMM_added/temporary/GPS2017.csv")
write_csv(GPS2018,"~/yodalab/BTG_Trj_U/analyze_data/middle_data/Added/No5_HMM_added/temporary/GPS2018.csv")
write_csv(GPS2019,"~/yodalab/BTG_Trj_U/analyze_data/middle_data/Added/No5_HMM_added/temporary/GPS2019.csv")
write_csv(GPS2020,"~/yodalab/BTG_Trj_U/analyze_data/middle_data/Added/No5_HMM_added/temporary/GPS2020.csv")
write_csv(GPS2021,"~/yodalab/BTG_Trj_U/analyze_data/middle_data/Added/No5_HMM_added/temporary/GPS2021.csv")
write_csv(GPS2022,"~/yodalab/BTG_Trj_U/analyze_data/middle_data/Added/No5_HMM_added/temporary/GPS2022.csv")
```


#18. Extract unclassified data (1st round)
Extract all IDs whose State column contains only NA values and combine
them into a single data frame. This process is repeated until all IDs
have been successfully classified.
```{r}
out_place <-"~/yodalab/BTG_Trj_U/analyze_data/middle_data/Added/No5_HMM_added/fixed/"
sampling_dt <- 300 

year_list <- c("GPS2013","GPS2014","GPS2015","GPS2016","GPS2017","GPS2018","GPS2019","GPS2020","GPS2021","GPS2022")

data_again <- tibble()
for (year in year_list) {
  
  final_data <- get(year)  
  grouped_data <- final_data %>% group_by(ID)
  NAtrip <- grouped_data %>% summarise(n_na = sum(is.na(State)))
  NAname <- paste0("NA", year)
  assign(NAname, NAtrip)
  #Extract the trip with NA > 50%
  Again <- final_data %>%
    group_by(ID) %>%
    summarize(na_ratio = sum(is.na(State)) / n()) %>%
    filter(na_ratio > 0.5) %>%
    ungroup()
  NAname <- paste0("delete", year)
  assign(NAname, Again)
  Again_matched <- semi_join(final_data, Again, by = "ID")
  data_again <- rbind(data_again, Again_matched)
}
```

#19. Split data into smaller chunks
The dataset is split because processing the full dataset at once
causes the computer to run out of memory.
```{r}
split_size <- nrow(data_again) %/% 2
data_again_1 <- data_again[1:split_size, ]
data_again_2 <- data_again[(split_size + 1):nrow(data_again), ]
```

#20. Chunk to run if an error occurs
HMM classification throws an error when the dataset contains
duplicate GPS points with identical latitude and longitude values.
If this occurs, run this chunk to remove the duplicates before
re-running the HMM classification above.
```{r}
remove_duplicates <- function(data_name) {
  data0 <- get(data_name)
  df <- data0[, c("x", "y")]
  num_duplicate_rows <- sum(duplicated(df))
  print(paste(data_name, "contains", num_duplicate_rows, "duplicates."))
  duplicate_rows <- data0 %>%
    group_by(x, y) %>%
    filter(n() > 1) %>% 
    arrange(x, y)
  data0 <- distinct(data0, x, y, .keep_all = TRUE)
  assign(data_name, data0, envir = .GlobalEnv)
  return(duplicate_rows)
}

data_names <- c("data_again_1","data_again_2") # Add the dataset name here to rerun

for (data_name in data_names) {
  remove_duplicates(data_name)
}
```

#21. HMM State Classification (2nd round)
```{r}
year_list <- c("data_again_1","data_again_2")
HMM_models_noduplication(year_list)
```

#22. 22. Reload GPS and HMM output
GPS2013 to GPS2022 contain the original GPS data with states assigned
by the 1st round of HMM classification (chunk 12).
The HMM files contain the re-classified results from the 2nd round
of HMM classification (chunk 21), applied only to IDs that could not
be classified in the 1st round.
```{r}
csv_files <- list.files(path = "~/yodalab/BTG_Trj_U/analyze_data/middle_data/Added/No5_HMM_added/temporary/", 
                        pattern = "\\.csv$", full.names = TRUE)
for (file in csv_files) {
  var_name <- tools::file_path_sans_ext(basename(file))
  assign(var_name, read_csv(file), envir = .GlobalEnv)
}

m3_1 <- read_csv("~/yodalab/BTG_Trj_U/analyze_data/middle_data/Added/No5_HMM_added/HMM_data_again_1.csv")
m3_2 <- read_csv("~/yodalab/BTG_Trj_U/analyze_data/middle_data/Added/No5_HMM_added/HMM_data_again_2.csv")

m3_again <- rbind(m3_1, m3_2)
```

#23. Match states to GPS data (2nd round)
```{r}
update_state <- function(year) {
  year_string <- as.character(year)
  data_name <- paste0(input_name, year_string)
  model_name <- "m3_again"
  
  data <- get(data_name)
  model <- get(model_name)

  for (i in 1:length(model$ID)) {
    index <- which(data$ID == model$ID[i] & data$time == model$time[i])
    data$State[index] <- model$State[i]
  }
  
  assign(data_name, data, envir = .GlobalEnv)
  
}

input_name <- "GPS"
years <- 2013:2022
lapply(years, update_state)
```

#24. Extract unclassified data (2nd round)
```{r}
final_data <- bind_rows(GPS2013,GPS2014,GPS2015,GPS2016,GPS2017,GPS2018,GPS2019,GPS2020,GPS2021,GPS2022)

#Extract the ID with NA > 80%
again_second <- final_data %>%
  group_by(ID) %>%
  filter(mean(is.na(State)) > 0.8) %>%
  ungroup()
```

#25. HMM State Classification (3rd round)
```{r}
year_list <- c("again_second")
HMM_models_noduplication(year_list)
```

#26. Match states to GPS data (3rd round)
```{r}
m3_second <- read_csv("~/yodalab/BTG_Trj_U/analyze_data/middle_data/Added/No5_HMM_added/HMM_again_second.csv")

update_state <- function(year) {
  year_string <- as.character(year)
  data_name <- paste0(input_name, year_string)
  model_name <- "m3_second"
  
  data <- get(data_name)
  model <- get(model_name)

  for (i in 1:length(model$ID)) {
    index <- which(data$ID == model$ID[i] & data$time == model$time[i])
    data$State[index] <- model$State[i]
  }
  
  assign(data_name, data, envir = .GlobalEnv)
  
}

input_name <- "GPS"
years <- 2013:2022
lapply(years, update_state)
```

#27. Extract unclassified data (3rd round)
```{r}
# 2. すべての更新されたデータフレームを結合
final_data <- bind_rows(GPS2013,GPS2014,GPS2015,GPS2016,GPS2017,GPS2018,GPS2019,GPS2020,GPS2021,GPS2022)

# 3. NAの割合が80%以上のIDを抽出して、そのIDを含むすべての行をdata_again_thirdに保存
again_third <- final_data %>%
  group_by(ID) %>%
  filter(mean(is.na(State)) > 0.8) %>%
  ungroup()
#again_thirdがcharacter(0)ならここでやめてOK
```


#28. HMM State Classification (4th round)
Repeat the run until classification succeeds.
Note: An error will occur if the dataset contains only one unique ID.
```{r}
year_list <- c("again_third")
HMM_models_noduplication(year_list)
```

#29. Match states to GPS data (4th round)
```{r}
m3_third <- read_csv("~/yodalab/BTG_Trj_U/analyze_data/middle_data/Added/No5_HMM_added/HMM_again_third.csv")

update_state <- function(year) {
  year_string <- as.character(year)
  data_name <- paste0(input_name, year_string)
  model_name <- "m3_third"
  
  data <- get(data_name)
  model <- get(model_name)

  for (i in 1:length(model$ID)) {
    index <- which(data$ID == model$ID[i] & data$time == model$time[i])
    data$State[index] <- model$State[i]
  }
  
  assign(data_name, data, envir = .GlobalEnv)
  
}

input_name <- "GPS"
years <- 2013:2022
lapply(years, update_state)
```

#30. Extract unclassified data (4th round)
```{r}
final_data <- bind_rows(GPS2013,GPS2014,GPS2015,GPS2016,GPS2017,GPS2018,GPS2019,GPS2020,GPS2021,GPS2022)
again_last <- final_data %>%
  group_by(ID) %>%
  filter(mean(is.na(State)) > 0.8) %>%
  ungroup()
#If final_data character(0): stop here
```

#31. Visualize
```{r}
my_palette <- c("cadetblue4", "yellowgreen", "#E69F00", "grey", "#0072B2", "#D55E00")

out_place <- "~/yodalab/BTG_Trj_U/analyze_data/middle_data/Added/No5_HMM_added/Trip_plot/"
year_list <- list("GPS2013","GPS2014","GPS2015","GPS2016","GPS2017","GPS2018","GPS2019","GPS2020","GPS2021","GPS2022")

japan_map <- map_data("world", region = "Japan")

x_range <- c(141.1705, 143.0207)
y_range <- c(38.82596, 41.48974)

for (year in year_list) {
  final_data <- get(year)
  id_list <- unique(final_data$ID)
  n_id <- length(id_list)

  for(i in 1:n_id){
    data1 <- final_data %>%
      filter(ID==id_list[i])

    plot <- ggplot() +
      geom_polygon(data = japan_map, aes(x = long, y = lat, group = group), fill = "white", color = "black") +
      geom_point(data = data1, aes(x = lon, y = lat, colour = State), size = 1) +
      geom_path(size=1.5, alpha=0.5, colour= "snow2") +
      coord_cartesian(xlim = x_range, ylim = y_range) +  # プロットの範囲を指定
      theme_void() +
      scale_color_manual(values = my_palette)

    ggsave(file = paste0(out_place,"map", id_list[i], ".png"), plot = plot)
  }
}

    plot <- ggplot() +
      geom_point(data = japan_map, aes(x = long, y = lat, group = group), fill = "white", color = "black")+
      coord_cartesian(xlim = x_range, ylim = y_range)
```

#32. Output
```{r}
write_csv(GPS2013,"~/yodalab/BTG_Trj_U/analyze_data/middle_data/Added/No5_HMM_added/fixed/GPS2013.csv")
write_csv(GPS2014,"~/yodalab/BTG_Trj_U/analyze_data/middle_data/Added/No5_HMM_added/fixed/GPS2014.csv")
write_csv(GPS2015,"~/yodalab/BTG_Trj_U/analyze_data/middle_data/Added/No5_HMM_added/fixed/GPS2015.csv")
write_csv(GPS2016,"~/yodalab/BTG_Trj_U/analyze_data/middle_data/Added/No5_HMM_added/fixed/GPS2016.csv")
write_csv(GPS2017,"~/yodalab/BTG_Trj_U/analyze_data/middle_data/Added/No5_HMM_added/fixed/GPS2017.csv")
write_csv(GPS2018,"~/yodalab/BTG_Trj_U/analyze_data/middle_data/Added/No5_HMM_added/fixed/GPS2018.csv")
write_csv(GPS2019,"~/yodalab/BTG_Trj_U/analyze_data/middle_data/Added/No5_HMM_added/fixed/GPS2019.csv")
write_csv(GPS2020,"~/yodalab/BTG_Trj_U/analyze_data/middle_data/Added/No5_HMM_added/fixed/GPS2020.csv")
write_csv(GPS2021,"~/yodalab/BTG_Trj_U/analyze_data/middle_data/Added/No5_HMM_added/fixed/GPS2021.csv")
write_csv(GPS2022,"~/yodalab/BTG_Trj_U/analyze_data/middle_data/Added/No5_HMM_added/fixed/GPS2022.csv")
```

#33. read data again
```{r}
rm(list=ls())

#data with removed coast and added the state classification
csv_files <- list.files(path = "~/yodalab/BTG_Trj_U/analyze_data/middle_data/Added/No5_HMM_added/fixed/", 
                        pattern = "\\.csv$", full.names = TRUE)
for (file in csv_files) {
  var_name <- tools::file_path_sans_ext(basename(file))
  assign(var_name, read_csv(file), envir = .GlobalEnv)
}
  
#data before removing coast
csv_files <- list.files(path = "~/yodalab/BTG_Trj_U/analyze_data/middle_data/Added/No5_coast_added/before/", 
                        pattern = "\\.csv$", full.names = TRUE)
for (file in csv_files) {
  var_name <- tools::file_path_sans_ext(basename(file))
  assign(var_name, read_csv(file), envir = .GlobalEnv)
}
```

#34. Match states to the original data
Match the fully classified data (GPS2013 to GPS2022) back to the
original GPS data that includes coastal points (Original2013 to Original2022).
```{r}
update_state <- function(year) {
  year_string <- as.character(year)
  data_name <- paste0(input_name, year_string)
  model_name <- paste0(state_name, year_string)
  
  data <- get(data_name)
  model <- get(model_name)
  
  data$State <- NA
  for (i in 1:length(model$ID)) {
    index <- which(data$ID == model$ID[i] & data$time == model$time[i])
    data$State[index] <- model$State[i]
  }
  
  assign(data_name, data, envir = .GlobalEnv)
  
}

input_name <- "Original"
state_name <- "GPS"
years <- 2013:2022
lapply(years, update_state)
```

#35. Remove GPS points on land
GPS points located on land are not used in this analysis.
Set their State values to NA.
```{r}
# Natural Earth data download
ocean <- ne_download(scale = 50, type = 'ocean', category = 'physical', returnclass = 'sf')

# function to define sea/land
classify_sealand <- function(df) {
  if (!all(c('lon', 'lat') %in% colnames(df))) {
    stop("lon or lat don't exist")
  }
  coords <- st_as_sf(df, coords = c("lon", "lat"), crs = st_crs(ocean))
  sealand_matrix <- ifelse(st_intersects(coords, ocean, sparse = FALSE), "Land", "Sea")

  df$sealand <- apply(sealand_matrix, 1, function(x) {
    if ("Land" %in% x) {
      return("Land")
    } else {
      return("Sea")
    }
  })

  return(df)
}

dfs <- list(Original2013, Original2014, Original2015, Original2016, Original2017, Original2018, Original2019, Original2020, Original2021, Original2022)

dfs_processed <- lapply(dfs, function(df) {
  df <- classify_sealand(df) %>%
    group_by(ID) %>%
    mutate(
      first_Fast_Homing = ifelse(any(State == "Fast_Homing"), min(which(State == "Fast_Homing")), NA),

      State = ifelse(is.na(first_Fast_Homing) | row_number() < first_Fast_Homing, State, 
                     ifelse(sealand == "Land", NA_character_, State))
    ) %>%
    ungroup() %>%
    dplyr::select(-first_Fast_Homing)

  return(df)
})

list2env(setNames(dfs_processed, c("Original2013", "Original2014", "Original2015", "Original2016", "Original2017", "Original2018", "Original2019", "Original2020", "Original2021", "Original2022")), envir = .GlobalEnv)
```

#36. Visualize
```{r}
my_palette <- c("cadetblue4", "yellowgreen", "#E69F00", "grey", "#0072B2", "#D55E00")

out_place <- "~/yodalab/BTG_Trj_U/analyze_data/middle_data/Added/No5_HMM_added/Trip_plot/"
year_list <- list("Original2013","Original2014", "Original2015", "Original2016", "Original2017", "Original2018", "Original2019","Original2020","Original2021","Original2022")

japan_map <- map_data("world", region = "Japan")

x_range <- c(141.1705, 143.0207)
y_range <- c(38.82596, 41.48974)

for (year in year_list) {
  final_data <- get(year)
  id_list <- unique(final_data$ID)
  n_id <- length(id_list)

  for(i in 1:n_id){
    data1 <- final_data %>%
      filter(ID==id_list[i])

    plot <- ggplot() +
      geom_polygon(data = japan_map, aes(x = long, y = lat, group = group), fill = "white", color = "black") +
      geom_point(data = data1, aes(x = lon, y = lat, colour = State), size = 1) +
      geom_path(size=1.5, alpha=0.5, colour= "snow2") +
      coord_cartesian(xlim = x_range, ylim = y_range) +  
      theme_void() +
      scale_color_manual(values = my_palette)

    ggsave(file = paste0(out_place,"map", id_list[i], ".png"), plot = plot)
  }
}

    plot <- ggplot() +
      geom_point(data = japan_map, aes(x = long, y = lat, group = group), fill = "white", color = "black")+
      coord_cartesian(xlim = x_range, ylim = y_range)
```

###################################################################
Check the output data and adjust the conditions as needed.
For example, try extracting exceptional trips individually and
re-running HMM without excluding coastal points.
###################################################################

#37. Output
```{r}
write_csv(Original2013,"~/yodalab/BTG_Trj_U/analyze_data/final_data/Everything/GPS2013.csv")
write_csv(Original2014,"~/yodalab/BTG_Trj_U/analyze_data/final_data/Everything/GPS2014.csv")
write_csv(Original2015,"~/yodalab/BTG_Trj_U/analyze_data/final_data/Everything/GPS2015.csv")
write_csv(Original2016,"~/yodalab/BTG_Trj_U/analyze_data/final_data/Everything/GPS2016.csv")
write_csv(Original2017,"~/yodalab/BTG_Trj_U/analyze_data/final_data/Everything/GPS2017.csv")
write_csv(Original2018,"~/yodalab/BTG_Trj_U/analyze_data/final_data/Everything/GPS2018.csv")
write_csv(Original2019,"~/yodalab/BTG_Trj_U/analyze_data/final_data/Everything/GPS2019.csv")
write_csv(Original2020,"~/yodalab/BTG_Trj_U/analyze_data/final_data/Everything/GPS2020.csv")
write_csv(Original2021,"~/yodalab/BTG_Trj_U/analyze_data/final_data/Everything/GPS2021.csv")
write_csv(Original2022,"~/yodalab/BTG_Trj_U/analyze_data/final_data/Everything/GPS2022.csv")
```