---
title: "Wind"
output: html_document
date: "2025-01-10"
---
In previous code (Code4 Wind), I used the function match_wind: Matches each GPS point to the single nearest ERA5 grid point in space and time. Simple and fast, but less accurate.

In here, I used conv_wind2: Applies inverse-distance-weighted interpolation across the four surrounding ERA5 grid points, with temporal interpolation between the two nearest time steps. Slower but more accurate.

#library
```{r}
library(dplyr)
library(lubridate)
library(ncdf4)
library(tidyverse)
library(geosphere)
library(rnaturalearth)
library(sf)
library(ggspatial)
```

#data
```{r}
rm(list=ls())
sea_data <- read_csv("~/Documents/yodalab/data/final_data/all_sea_data.csv")
```


#Check the range of GPS and wind
```{r}
check_data_range <- function(data, target_year, wind_dir = "~/Documents/yodalab/data/wind/") {
  wind_file <- paste0(wind_dir, "wind_", target_year, ".nc")
  nc_data <- nc_open(wind_file)
  
  wind_time <- as.POSIXct(ncvar_get(nc_data, "valid_time"), origin="1970-01-01", tz="UTC")
  wind_lon <- ncvar_get(nc_data, "longitude")
  wind_lat <- ncvar_get(nc_data, "latitude")

  cat(sprintf("\n%d year wind data：\n", target_year))
  cat("time range：", format(min(wind_time)), "to", format(max(wind_time)), "\n")
  cat("lon range：", min(wind_lon), "to", max(wind_lon), "\n")
  cat("lat range：", min(wind_lat), "to", max(wind_lat), "\n")
  
  track_year <- data %>% 
    filter(year == target_year)
  
  cat(sprintf("\n%d year tracking data：\n", target_year))
  if (nrow(track_year) > 0) {
    cat("time range：", format(min(track_year$time)), "to", format(max(track_year$time)), "\n")
    cat("lon range：", min(track_year$lon), "to", max(track_year$lon), "\n")
    cat("lat range：", min(track_year$lat), "to", max(track_year$lat), "\n")
    cat("data point：", nrow(track_year), "\n")
  } else {
    cat("no tracking data\n")
  }
  nc_close(nc_data)
  return(list(
    wind = list(
      time_range = range(wind_time),
      lon_range = range(wind_lon),
      lat_range = range(wind_lat)
    ),
    track = if(nrow(track_year) > 0) list(
      time_range = range(track_year$time),
      lon_range = range(track_year$lon),
      lat_range = range(track_year$lat),
      n_points = nrow(track_year)
    ) else NULL
  ))
}

years_to_check <- 2013:2022
for(year in years_to_check) {
  tryCatch({
    check_data_range(sea_data, year)
  }, error = function(e) {
    cat(sprintf("\n%dyear error：%s\n", year, e$message))
  })
}

```

#Add wind
Wind is interpolated to each GPS point using inverse distance weighting
across the 4 surrounding ERA5 grid points (spatial) and the 2 bracketing time steps (temporal), giving a more accurate estimate than nearest-neighbor matching.
```{r}
# # Remove ERA5 wind columns from sea_data
# sea_data <- sea_data %>%
#   dplyr::select(-ERA5_wind_east, -ERA5_wind_north)

conv_wind2 <- function(nc_data, data_time, data_lon, data_lat) {
 if (!identical(attr(data_time, "tzone"), "UTC")) {
   warning("Input data_time is not in UTC. Converting to UTC...")
   data_time <- with_tz(data_time, "UTC")
 }
 
 lon <- ncdf4::ncvar_get(nc_data, "longitude")
 lat <- ncdf4::ncvar_get(nc_data, "latitude")
 time <- ncdf4::ncvar_get(nc_data, "valid_time")
 time_obs <- as.POSIXct(time, origin="1970-01-01", tz="UTC")
 
 cat(sprintf("Data time range: %s to %s (UTC)\n", 
             min(data_time), max(data_time)))
 cat(sprintf("NC file time range: %s to %s (UTC)\n", 
             min(time_obs), max(time_obs)))
 
 if (
   (min(data_time, na.rm = TRUE) < min(time_obs, na.rm = TRUE)) ||
   (max(time_obs, na.rm = TRUE) < max(data_time, na.rm = TRUE)) ||
   (min(data_lon, na.rm = TRUE) < min(lon, na.rm = TRUE)) ||
   (max(lon, na.rm = TRUE) < max(data_lon, na.rm = TRUE)) ||
   (min(data_lat, na.rm = TRUE) < min(lat, na.rm = TRUE)) ||
   (max(lat) < max(data_lat, na.rm = TRUE))
 ) {
   print("The track data include data points that are outside the spatiotemporal range of the input environmental data.")
   return(NA)
 }
 
 u <- ncdf4::ncvar_get(nc_data, "u10")
 v <- ncdf4::ncvar_get(nc_data, "v10")
 
 n_lat <- length(lat)
 n_lon <- length(lon)
 n <- length(data_time)
 
 ERA5_wind_east <- rep(NA, n)
 ERA5_wind_north <- rep(NA, n)
 
 Time_low <- rep(as.POSIXct(NA), n)
 Time_up <- rep(as.POSIXct(NA), n)
 Lat_low <- rep(NA, n)
 Lat_up <- rep(NA, n)
 Lon_low <- rep(NA, n)
 Lon_up <- rep(NA, n)
 
 pb <- txtProgressBar(min = 0, max = n, style = 3)
 
 for(i in 1:n) {
   setTxtProgressBar(pb, i)
   
   diff_vector <- lat - data_lat[i]
   lat_up_id <- which(diff_vector == min(diff_vector[diff_vector >= 0]))
   lat_low_id <- which(diff_vector == max(diff_vector[diff_vector < 0]))
   
   diff_vector <- lon - data_lon[i]
   lon_up_id <- which(diff_vector == min(diff_vector[diff_vector >= 0]))
   lon_low_id <- which(diff_vector == max(diff_vector[diff_vector < 0]))
   
   diff_vector <- as.numeric(time_obs - data_time[i])
   t_up_id <- which(diff_vector == min(diff_vector[diff_vector >= 0]))
   t_low_id <- which(diff_vector == max(diff_vector[diff_vector < 0]))
   
   Lat_up[i] <- lat[lat_up_id]
   Lat_low[i] <- lat[lat_low_id]
   Lon_up[i] <- lon[lon_up_id]
   Lon_low[i] <- lon[lon_low_id]
   Time_up[i] <- time_obs[t_up_id]
   Time_low[i] <- time_obs[t_low_id]
   
   d1 <- geosphere::distHaversine(c(data_lon[i], data_lat[i]),
                                c(Lon_low[i], Lat_up[i])) / 1000
   d2 <- geosphere::distHaversine(c(data_lon[i], data_lat[i]),
                                c(Lon_up[i], Lat_up[i])) / 1000
   d3 <- geosphere::distHaversine(c(data_lon[i], data_lat[i]),
                                c(Lon_low[i], Lat_low[i])) / 1000
   d4 <- geosphere::distHaversine(c(data_lon[i], data_lat[i]),
                                c(Lon_up[i], Lat_low[i])) / 1000
   
   ta <- as.numeric(data_time[i] - Time_low[i], units="hours")
   tb <- as.numeric(Time_up[i] - data_time[i], units="hours")
   
   calculate_spatial_interpolation <- function(u_or_v, d1, d2, d3, d4,
                                            lon_low_id, lon_up_id,
                                            lat_low_id, lat_up_id, time_id) {
     value_1 <- u_or_v[lon_low_id, lat_up_id, time_id]
     value_2 <- u_or_v[lon_up_id, lat_up_id, time_id]
     value_3 <- u_or_v[lon_low_id, lat_low_id, time_id]
     value_4 <- u_or_v[lon_up_id, lat_low_id, time_id]
     
     inverse_distance_weighted_mean <- (
       1/d1*value_1 + 1/d2*value_2 + 1/d3*value_3 + 1/d4*value_4
     ) / (1/d1 + 1/d2 + 1/d3 + 1/d4)
     
     return(inverse_distance_weighted_mean)
   }
   
   calculate_temporal_interpolation <- function(a, b, ta, tb) {
     if (ta == 0) {
       return(a)
     } else if (tb == 0) {
       return(b)
     } else {
       return((1/ta * a + 1/tb * b) / (1/ta + 1/tb))
     }
   }
   
   ERA5_wind_east_a_inv_mean <- calculate_spatial_interpolation(
     u, d1, d2, d3, d4, lon_low_id, lon_up_id, lat_low_id, lat_up_id, t_low_id
   )
   ERA5_wind_east_b_inv_mean <- calculate_spatial_interpolation(
     u, d1, d2, d3, d4, lon_low_id, lon_up_id, lat_low_id, lat_up_id, t_up_id
   )
   ERA5_wind_east[i] <- calculate_temporal_interpolation(
     ERA5_wind_east_a_inv_mean, ERA5_wind_east_b_inv_mean, ta, tb
   )
   
   ERA5_wind_north_a_inv_mean <- calculate_spatial_interpolation(
     v, d1, d2, d3, d4, lon_low_id, lon_up_id, lat_low_id, lat_up_id, t_low_id
   )
   ERA5_wind_north_b_inv_mean <- calculate_spatial_interpolation(
     v, d1, d2, d3, d4, lon_low_id, lon_up_id, lat_low_id, lat_up_id, t_up_id
   )
   ERA5_wind_north[i] <- calculate_temporal_interpolation(
     ERA5_wind_north_a_inv_mean, ERA5_wind_north_b_inv_mean, ta, tb
   )
 }
 
 close(pb)
 
 result <- data.frame(ERA5_wind_east, ERA5_wind_north)
 return(result)
}

process_wind_data <- function(data, wind_dir = "~/Documents/yodalab/data/wind/") {
 if (!identical(attr(data$time, "tzone"), "UTC")) {
   warning("Input data time is not in UTC. Converting to UTC...")
   data$time <- with_tz(data$time, "UTC")
 }
 
 data_by_year <- data %>%
   group_by(year) %>%
   group_split()
 
 results_list <- list()
 
 for (year_data in data_by_year) {
   current_year <- unique(year_data$year)
   cat(sprintf("Processing year %d...\n", current_year))
   
   wind_file <- paste0(wind_dir, "wind_", current_year, ".nc")
   
   nc_data <- ncdf4::nc_open(wind_file)
   
   wind_result <- conv_wind2(
     nc_data = nc_data,
     data_time = year_data$time,
     data_lon = year_data$lon,
     data_lat = year_data$lat
   )
   
   if (!all(is.na(wind_result))) {
     year_result <- cbind(year_data, wind_result)
     
     results_list[[as.character(current_year)]] <- year_result
   }
   
   ncdf4::nc_close(nc_data)
 }
 
 final_result <- bind_rows(results_list)
 return(final_result)
}

cat("Checking timezone of input data:\n")
cat("sea_data timezone:", attr(sea_data$time, "tzone"), "\n")

sea_data_with_wind <- process_wind_data(sea_data)
```

#Calculate backwind, crosswind,α
Reference: https://www.ncbi.nlm.nih.gov/pmc/articles/PMC7661306/
```{r}
data_all <- sea_data_with_wind

data_out <- tibble()
ids <- unique(data_all$HomingID)
nest_lon <- 141.5576031
nest_lat <- 40.5386062

deg2rad <- function(deg) {
  return(deg * pi / 180)
}

for (selectid in ids) {
  data <- filter(data_all, HomingID == selectid)

# Compute the direction to the nest
  data <- data %>%
    mutate(direction_to_nest = bearing(cbind(lon, lat), c(nest_lon, nest_lat)))

# Compute the direction of movement
  data <- data %>%
    mutate(direction_of_movement = bearing(cbind(lon, lat), cbind(lead(lon), lead(lat))))
  
# Compute α
data <- data %>%
  mutate(
    α = direction_of_movement - direction_to_nest
  )

data <- data %>%
  mutate(
    α = ifelse(α > 180, α - 360, ifelse(α < -180, α + 360, α))
  )
    
data <- data %>%
  mutate(
    direction_to_nest = ifelse(direction_to_nest < 0, direction_to_nest + 360, direction_to_nest),
    direction_of_movement = ifelse(direction_of_movement < 0, direction_of_movement + 360, direction_of_movement)
  )

data <- data %>%
  rename(
    Wx = ERA5_wind_east,
    Wy = ERA5_wind_north
  )

  data <- data %>%
    mutate(
      nx = sin(deg2rad(direction_to_nest)), # Unit vector toward the nest (east component)
      ny = cos(deg2rad(direction_to_nest)), # Unit vector toward the nest (north component)
      wind_back = Wx * nx + Wy * ny,  # Tailwind component along the beeline to the nest
      wind_cross = Wx * ny - Wy * nx   # Crosswind component perpendicular to the beeline
    )
  
  data_out <- data_out %>% bind_rows(data)
}

write_csv(data_out, "~/Documents/yodalab/data/final_data/all_sea_data_wind.csv")
```
