---
title: "HMM2"
output: html_document
date: "2023-09-12"
---
HMM classification cannot classify all IDs in a single run.
It is necessary to run the classification multiple times and verify
that all IDs have been successfully assigned a state.
Run this script as a continuation of New5_HMM1.Rmd.

Back to Code2A_HMM1.Rmd's #8. HMM State Classification and # 7. Function to remove consecutive duplicate GPS points Chunk if you accidentally removed the function

#2nd round
#14. read GPS data before running HMM
```{r}
#14. read GPS data before running HMM

years <- c(2013)
# years <- 2013:2022

base_path <- "~/Documents/yodalab/paper/JAB/data/not_for_upload/"
hmm_path <- paste0(base_path, "HMM_added/")
input_path <- paste0(hmm_path, "input/")
processed_path <- paste0(hmm_path, "processed/")
unprocessed_path <- paste0(hmm_path, "unprocessed/")
output_path <- paste0(hmm_path, "output/")
trip_path <- paste0(hmm_path, "trip/")

dir.create(processed_path, showWarnings = FALSE, recursive = TRUE)
dir.create(unprocessed_path, showWarnings = FALSE, recursive = TRUE)
dir.create(output_path, showWarnings = FALSE, recursive = TRUE)
dir.create(trip_path, showWarnings = FALSE, recursive = TRUE)

read_gps_year <- function(yr) {
  
  gps_files <- list.files(
    path = input_path,
    pattern = paste0("^GPS", yr, "(_[0-9]+)?\\.csv$"),
    full.names = TRUE
  )
  
  gps_files <- sort(gps_files)
  
  if (length(gps_files) == 0) {
    stop(paste0("No GPS files found for ", yr))
  }
  
  message("Reading GPS files for ", yr, ": ",
          paste(basename(gps_files), collapse = ", "))
  
  read_csv(gps_files, show_col_types = FALSE)
}

for (yr in years) {
  assign(
    paste0("GPS", yr),
    read_gps_year(yr),
    envir = .GlobalEnv
  )
}
```

#15. read HMM classified data
```{r}
#15. read HMM classified data

read_hmm_year <- function(yr) {
  
  hmm_files <- list.files(
    path = hmm_path,
    pattern = paste0("^HMM_GPS", yr, "(_[0-9]+)?\\.csv$"),
    full.names = TRUE
  )
  
  hmm_files <- sort(hmm_files)
  
  if (length(hmm_files) == 0) {
    stop(paste0("No HMM files found for ", yr))
  }
  
  message("Reading HMM files for ", yr, ": ",
          paste(basename(hmm_files), collapse = ", "))
  
  read_csv(hmm_files, show_col_types = FALSE)
}

for (yr in years) {
  assign(
    paste0("HMM", yr),
    read_hmm_year(yr),
    envir = .GlobalEnv
  )
}
```

#16. Match states to GPS data
After this step, the State column for each ID will contain:
  - Success: one of the 4 states (Fast_Outbound, Fast_Middle,
             Fast_Homing, Slow), with NA only at the last point
  - Failure: NA for all points
```{r}
update_state <- function(gps_df, hmm_df) {
  
  # If GPS data does not have State column yet, create it
  if (!"State" %in% colnames(gps_df)) {
    gps_df$State <- NA_character_
  }
  
  # Prepare HMM state data
  # Use time_key to avoid problems caused by different time formats
  hmm_state <- hmm_df %>%
    dplyr::select(ID, time, State) %>%
    mutate(time_key = as.character(time)) %>%
    dplyr::select(ID, time_key, State_new = State) %>%
    distinct(ID, time_key, .keep_all = TRUE)
  
  # Keep existing State, and replace it only when new HMM State exists
  gps_updated <- gps_df %>%
    mutate(time_key = as.character(time)) %>%
    left_join(hmm_state, by = c("ID", "time_key")) %>%
    mutate(
      State = ifelse(
        !is.na(State_new),
        State_new,
        State
      )
    ) %>%
    dplyr::select(-time_key, -State_new)
  
  return(gps_updated)
}

for (yr in years) {
  
  gps_name <- paste0("GPS", yr)
  hmm_name <- paste0("HMM", yr)
  
  gps_df <- get(gps_name)
  hmm_df <- get(hmm_name)
  
  updated_df <- update_state(gps_df, hmm_df)
  
  assign(gps_name, updated_df, envir = .GlobalEnv)
}
```

#17. Output
In case your laptop will die
```{r}
for (yr in years) {
  gps_name <- paste0("GPS", yr)
  
  write_csv(
    get(gps_name),
    paste0(processed_path, gps_name, ".csv")
  )
}
```


#18 Define the function extract_unclassified_data
Extract all IDs whose State column contains only NA values and combine
them into a single data frame. This process is repeated until all IDs
have been successfully classified.
```{r}
extract_unclassified_data <- function(years, threshold, output_name) {
  
  unclassified_data <- tibble()
  
  for (yr in years) {
    
    gps_name <- paste0("GPS", yr)
    final_data <- get(gps_name)
    
    # Check IDs where all State values are NA
    all_na_id <- final_data %>%
      group_by(ID) %>%
      summarise(
        all_na = all(is.na(State)),
        .groups = "drop"
      ) %>%
      filter(all_na)
    
    message("IDs with all State = NA in ", gps_name, ":")
    print(all_na_id)
    
    # Extract IDs where the proportion of NA State is above the threshold
    again_id <- final_data %>%
      group_by(ID) %>%
      summarise(
        na_ratio = mean(is.na(State)),
        .groups = "drop"
      ) %>%
      filter(na_ratio > threshold)
    
    again_data <- semi_join(final_data, again_id, by = "ID") %>%
      dplyr::select(-any_of("State"))
    
    unclassified_data <- bind_rows(unclassified_data, again_data)
  }
  
  write_csv(
    unclassified_data,
    paste0(unprocessed_path, output_name, ".csv")
  )
  
  write_csv(
    unclassified_data,
    paste0(input_path, output_name, ".csv")
  )
  
  return(unclassified_data)
}
```

#19 Extract unclassified data
```{r}
data_again <- extract_unclassified_data(
  years = years,
  threshold = 0.5,
  output_name = "data_again"
)
```

#20. Split data into smaller chunks
The dataset is split because processing the full dataset at once
causes the computer to run out of memory.
```{r}
if (nrow(data_again) == 0) {
  stop("All IDs were classified. No need to run HMM again.")
}

split_size <- nrow(data_again) %/% 2

data_again_1 <- data_again[1:split_size, ]
data_again_2 <- data_again[(split_size + 1):nrow(data_again), ]

write_csv(data_again_1, paste0(input_path, "data_again_1.csv"))
write_csv(data_again_2, paste0(input_path, "data_again_2.csv"))
```



#21. HMM State Classification (2nd round)
```{r}
year_list <- c("data_again_1","data_again_2")
HMM_models_noduplication(year_list)
```

#3rd round
#22. Reload GPS and HMM output
GPS2013 to GPS2022 contain the original GPS data with states assigned
by the 1st round of HMM classification (chunk 12).
The HMM files contain the re-classified results from the 2nd round
of HMM classification (chunk 21), applied only to IDs that could not
be classified in the 1st round.
```{r}
# Read processed GPS files (the one added state by the 1st round)
for (yr in years) {
  
  gps_name <- paste0("GPS", yr)
  gps_file <- paste0(processed_path, gps_name, ".csv")
  
  assign(
    gps_name,
    read_csv(gps_file, show_col_types = FALSE),
    envir = .GlobalEnv
  )
}

# Read HMM files from the 2nd round
again_files <- list.files(
  path = hmm_path,
  pattern = "^HMM_data_again_.*\\.csv$",
  full.names = TRUE
)

again_files <- sort(again_files)

message(
  "Reading 2nd-round HMM files: ",
  paste(basename(again_files), collapse = ", ")
)

m3_again <- read_csv(
  again_files,
  show_col_types = FALSE
)
```

#23. Match states to GPS data
```{r}
for (yr in years) {
  
  gps_name <- paste0("GPS", yr)
  gps_df <- get(gps_name)
  
  updated_df <- update_state(gps_df, m3_again)
  
  assign(gps_name, updated_df, envir = .GlobalEnv)
}
```

#24. Extract unclassified data
```{r}
#24. Extract unclassified data after 2nd round

again_second <- extract_unclassified_data(
  years = years,
  threshold = 0.5,
  output_name = "again_second"
)

if (nrow(again_second) == 0) {
  message("All IDs are classified enough. No need to run the 3rd round.")
} else {
  message("Some IDs still need re-classification. Run the 3rd round.")
}
```


#25. HMM State Classification (3rd round)
```{r}
if (nrow(again_second) > 0) {
  year_list <- c("again_second")
  HMM_models_noduplication(year_list)
} else {
  message("All IDs were classified after the 2nd round.")
}
```

#26. Match states to GPS data (3rd round)
```{r}
#26. Match states to GPS data (3rd round)

third_round_file <- paste0(hmm_path, "HMM_again_second.csv")

if (file.exists(third_round_file)) {
  
  # Read HMM output from the 3rd round
  m3_second <- read_csv(
    third_round_file,
    show_col_types = FALSE
  )
  
  # Update GPS data using the new HMM states
  for (yr in years) {
    
    gps_name <- paste0("GPS", yr)
    gps_df <- get(gps_name)
    
    updated_df <- update_state(gps_df, m3_second)
    
    assign(gps_name, updated_df, envir = .GlobalEnv)
  }
  
} else {
  message("No 3rd-round HMM output found. Skipping update.")
}
```

#27. Visualize
```{r}
my_palette <- c(
  "Fast_Outbound" = "#E69F00",
  "Fast_Middle"   = "yellowgreen",
  "Fast_Homing"   = "cadetblue4",
  "Slow"          = "grey"
)

japan_map <- map_data("world", region = "Japan")

x_range <- c(141.1705, 143.0207)
y_range <- c(38.82596, 41.48974)

for (yr in years) {
  
  gps_name <- paste0("GPS", yr)
  final_data <- get(gps_name)
  id_list <- unique(final_data$ID)
  
  for (i in seq_along(id_list)) {
    
    data1 <- final_data %>%
      filter(ID == id_list[i])
    
    p <- ggplot() +
      geom_polygon(
        data = japan_map,
        aes(x = long, y = lat, group = group),
        fill = "white",
        color = "black"
      ) +
      geom_path(
        data = data1,
        aes(x = lon, y = lat),
        linewidth = 1.2,
        alpha = 0.5,
        color = "snow2"
      ) +
      geom_point(
        data = data1,
        aes(x = lon, y = lat, color = State),
        size = 1
      ) +
      coord_cartesian(xlim = x_range, ylim = y_range) +
      theme_void() +
      scale_color_manual(values = my_palette)
    
    ggsave(
      filename = paste0(trip_path, "map_", id_list[i], ".png"),
      plot = p,
      width = 6,
      height = 5,
      dpi = 300
    )
  }
}
```

#28. Output
```{r}
for (yr in years) {
  
  gps_name <- paste0("GPS", yr)
  
  write_csv(
    get(gps_name),
    paste0(output_path, gps_name, ".csv")
  )
}
```

#29. read data again
```{r}
for (yr in years) {
  
  gps_name <- paste0("GPS", yr)
  gps_file <- paste0(output_path, gps_name, ".csv")
  
  assign(
    gps_name,
    read_csv(gps_file, show_col_types = FALSE),
    envir = .GlobalEnv
  )
}
```

#30. Remove GPS points on land
GPS points located on land are not used in this analysis.
Set their State values to NA.
```{r}
#30. Remove GPS points on land
# GPS points located on land are not used in this analysis.
# Set their State values to NA.

# Download Natural Earth ocean data
ocean <- ne_download(
  scale = 50,
  type = "ocean",
  category = "physical",
  returnclass = "sf"
)

# Function to classify GPS points as sea or land
classify_sealand <- function(df) {
  
  # Check whether lon and lat columns exist
  if (!all(c("lon", "lat") %in% colnames(df))) {
    stop("lon or lat column does not exist.")
  }
  
  coords <- st_as_sf(
    df,
    coords = c("lon", "lat"),
    crs = st_crs(ocean)
  )
  
  sealand_matrix <- ifelse(
    st_intersects(coords, ocean, sparse = FALSE),
    "Land",
    "Sea"
  )
  
  # Convert sealand_matrix into a single character vector
  df$sealand <- apply(sealand_matrix, 1, function(x) {
    if ("Land" %in% x) {
      return("Land")
    } else {
      return("Sea")
    }
  })
  
  return(df)
}

# Make a list of GPS data frames
dfs <- mget(paste0("GPS", years))

# Process each GPS data frame
dfs_processed <- lapply(dfs, function(df) {
  
  df <- classify_sealand(df) %>%
    group_by(ID) %>%
    mutate(
      first_Fast_Homing = ifelse(
        any(State == "Fast_Homing", na.rm = TRUE),
        min(which(State == "Fast_Homing")),
        NA_integer_
      ),
      
      # After the first Fast_Homing point,
      # set State to NA if the point is classified as Land
      State = ifelse(
        is.na(first_Fast_Homing) | row_number() < first_Fast_Homing,
        State,
        ifelse(sealand == "Land", NA_character_, State)
      )
    ) %>%
    ungroup() %>%
    dplyr::select(-first_Fast_Homing)
  
  return(df)
})

# Reassign processed data frames back to GPS2013, GPS2014, etc.
names(dfs_processed) <- paste0("GPS", years)
list2env(dfs_processed, envir = .GlobalEnv)
```

#31. Visualize
```{r}
#36. Visualize after land removal

my_palette <- c(
  "Fast_Outbound" = "#E69F00",
  "Fast_Middle"   = "yellowgreen",
  "Fast_Homing"   = "cadetblue4",
  "Slow"          = "grey"
)

japan_map <- map_data("world", region = "Japan")

x_range <- c(141.1705, 143.0207)
y_range <- c(38.82596, 41.48974)

for (yr in years) {
  
  gps_name <- paste0("GPS", yr)
  final_data <- get(gps_name)
  id_list <- unique(final_data$ID)
  
  for (i in seq_along(id_list)) {
    
    data1 <- final_data %>%
      filter(ID == id_list[i])
    
    p <- ggplot() +
      geom_polygon(
        data = japan_map,
        aes(x = long, y = lat, group = group),
        fill = "white",
        color = "black"
      ) +
      geom_path(
        data = data1,
        aes(x = lon, y = lat),
        linewidth = 1.2,
        alpha = 0.5,
        color = "snow2"
      ) +
      geom_point(
        data = data1,
        aes(x = lon, y = lat, color = State),
        size = 1
      ) +
      coord_cartesian(xlim = x_range, ylim = y_range) +
      theme_void() +
      scale_color_manual(values = my_palette)
    
    ggsave(
      filename = paste0(trip_path, "Sea/", "map_land_removed_", id_list[i], ".png"),
      plot = p,
      width = 6,
      height = 5,
      dpi = 300
    )
  }
}
```

###################################################################
Check the output data and adjust the conditions as needed.
For example, try extracting exceptional trips individually and
re-running HMM without excluding coastal points.
###################################################################

#32. Final Output
```{r}
for (yr in years) {
  
  gps_name <- paste0("GPS", yr)
  
  write_csv(
    get(gps_name),
    paste0(output_path, gps_name, ".csv")
  )
}
```