---
title: "New4_Wind"
author: "Serina"
date: "2023-07-28"
output: html_document
---
Caution: Kept for formatting purposes; in practice, wind columns are removed and re-added after HMM fitting
Wind data is from ERA5α: https://cds.climate.copernicus.eu/datasets/reanalysis-era5-single-levels?tab=download

#library
```{r}
rm(list=ls())
library(ncdf4)
library(tidyverse)
library(ggplot2)
library(geosphere)
library(gridExtra)
```

#data
```{r}
rm(list=ls())
csv_files <- list.files(path = "~/Documents/yodalab/data/middle_data/Added/No3_5min_resampled/fixed/", 
                        pattern = "\\.csv$", full.names = TRUE)
for (file in csv_files) {
  var_name <- tools::file_path_sans_ext(basename(file))
  assign(var_name, read_csv(file), envir = .GlobalEnv)
}
```

#chack the range of longitude/latitude
Download the wind data based on this range
```{r}
dataframes <- c("GPS2013","GPS2014","GPS2015","GPS2016","GPS2017","GPS2018","GPS2019","GPS2020","GPS2021","GPS2022")
result <- data.frame(Year = character(), LonMin = numeric(), LonMax = numeric(), LatMin = numeric(), LatMax = numeric(), TimeMin = character(), TimeMax = character())
for(df_name in dataframes) {
  df <- get(df_name)
  lon_min <- min(df$lon, na.rm = TRUE)
  lon_max <- max(df$lon, na.rm = TRUE)
  lat_min <- min(df$lat, na.rm = TRUE)
  lat_max <- max(df$lat, na.rm = TRUE)

  time_min <- min(df$time, na.rm = TRUE)
  time_max <- max(df$time, na.rm = TRUE)

  result <- rbind(result, data.frame(Year = df_name, LonMin = lon_min, LonMax = lon_max, LatMin = lat_min, LatMax = lat_max, TimeMin = as.character(time_min), TimeMax = as.character(time_max)))
}
print(result)
```

#Define the function match_wind
 A custom function to match ERA5 wind data to GPS tracking data based on time and location

 Arguments:
   nc_data    : NetCDF data containing ERA5 wind fields
   data_time  : Timestamp of the GPS data point
   data_lon   : Longitude of the GPS data point
   data_lat   : Latitude of the GPS data point

 For each GPS data point, the function selects the ERA5 grid point that is closest in time, longitude, and latitude, and returns a data frame with:
   ERA5_wind_east    : Eastward wind component at the selected ERA5 grid point
   ERA5_wind_north   : Northward wind component at the selected ERA5 grid point
   Time_diff_in_hours: Time difference between the GPS point and the matched ERA5 data
   Distance_in_km    : Spatial distance between the GPS point and the matched ERA5 grid point
```{r}
match_wind <- function(nc_data,data_time,data_lon,data_lat){
  lon  <- ncvar_get(nc_data,"longitude")
  lat  <- ncvar_get(nc_data,"latitude")

  #time in ERA5 data is "hours since 1900-01-01 00:00:00.0", so need to convert to POSIXct
  #If you want to check, you can use the code below
  #tunits <- ncatt_get(nc_data,"time","units")
  
  #(For old format)
  #time <- ncvar_get(nc_data,"time")
  #For new format
  time <- ncvar_get(nc_data,"valid_time")
  

  #(For old format) 
  # Time unit was "hours since 1900-01-01" (POSIX origin), so need to change second
  #time_obs <-as.POSIXct(time*3600,origin="1900-01-01 00:00:00",tz="GMT")
  #For new format
  # Time unit is "seconds since 1970-01-01" (POSIX origin), so no multiplication needed
  time_obs <- as.POSIXct(time, origin = "1970-01-01 00:00:00", tz = "GMT")
  
  
  #If you use the data other than ERA5, need to check the unit and convert time_obs by different manner
  
  #Eastward portion of wind vector (u10) and northward portion (v10)
  u <- ncvar_get(nc_data,"u10")
  v <- ncvar_get(nc_data,"v10")
  
  #convert missing data to NA
  fillvalue_u <- ncatt_get(nc_data,"u10","_FillValue")
  fillvalue_v <- ncatt_get(nc_data,"v10","_FillValue")
  u[u==fillvalue_u$value] <- NA
  v[v==fillvalue_v$value] <- NA
  
  
  # Define a function that returns the index of the element in vector y
  # closest to x_elem
find_closest_position <- function(x_elem, y) {
  differences <- abs(y - x_elem)
  return(which.min(differences))
}
# Find the index of the closest time in ERA5 for each observation time
   time_id <- sapply(data_time, find_closest_position, y = time_obs)
# Find the index of the closest longitude in ERA5 for each observation longitude
  lon_id  <- sapply(data_lon, find_closest_position, y = lon)
 # Find the index of the closest latitude in ERA5 for each observation latitude
  lat_id  <- sapply(data_lat, find_closest_position, y = lat)
#Chose the optimal point
  n <- length(time_id)
  ERA5_wind_east     <- rep(NA,n)
  ERA5_wind_north    <- rep(NA,n)
  Time_diff_in_hours <- rep(NA,n)
  Distance_in_km     <- rep(NA,n)
  
  for(i in 1:n){
    ERA5_wind_east[i]  <- u[lon_id[i], lat_id[i], time_id[i]]
    ERA5_wind_north[i] <- v[lon_id[i], lat_id[i], time_id[i]]
  Time_diff_in_hours[i] <- as.numeric(time_obs[time_id[i]] - data_time[i],unit="hours")
  Distance_in_km[i] <- geosphere::distGeo(c(data_lon[i],data_lat[i]),c(lon[lon_id[i]],lat[lat_id[i]]))/1000
  } 
  
  return(data.frame(ERA5_wind_east, 
                    ERA5_wind_north,
                    Time_diff_in_hours,
                    Distance_in_km
                    )
         )
  
}
```

#Add wind
```{r}
file_path <- "~/Documents/yodalab/data/middle_data/Added/No4_wind_added/"
data_latlon <- list(GPS2013, GPS2014, GPS2015, GPS2016, GPS2017, GPS2018, GPS2019, GPS2020, GPS2021, GPS2022)

file_names <- paste0("wind_",2013:2022, ".nc")

for (i in seq_along(file_names)) {
  
  nc_data <- nc_open(file.path("~/Documents/yodalab/data/wind", file_names[i]))
  
  # matchwind
  wind_ERA5 <- match_wind(nc_data,
                          data_time = data_latlon[[i]]$time,
                          data_lon = data_latlon[[i]]$lon,
                          data_lat = data_latlon[[i]]$lat)
  
  wind_data <- cbind(data_latlon[[i]], wind_ERA5)
  output_file <- paste0(file_path, "wind_data_", 2013+i-1, ".csv")
  write.csv(wind_data, output_file, row.names = FALSE)
}
```