---
title: "Code2 Outlier removal and trip segmentation"
author: "Serina"
date: "2023-07-26"
output: html_document
---
##library
```{r}
library(readr)
library(dplyr)
library(lubridate)
library(stringr)
library(SDLfilter)
library(tidyverse)
library(ggplot2)
library(sp)
library(crawl)
library(xts)
library(geosphere)
library(adehabitatLT)
library(tidyr)
library(ncdf4)
library(accelerometry)
library(gridExtra)
library(momentuHMM)
library(tictoc)
library(osmdata)
library(sf)
library(ggmap)
library(leaflet)
```

##file
```{r}
rm(list=ls())
csv_files <- list.files(path = "~/Documents/yodalab/data/middle_data/GPS_cleaned/", 
                        pattern = "\\.csv$", full.names = TRUE)

for (file in csv_files) {
  # ファイル名（拡張子除く）を変数名とする
  var_name <- tools::file_path_sans_ext(basename(file))
  
  # assign関数を使って、変数名で指定した名前の変数にデータを代入する
  assign(var_name, read_csv(file), envir = .GlobalEnv)
}

####idファイル読み込み
# フォルダ内のCSVファイル名のリストを取得する
csv_files <- list.files(path = "~/Documents/BTG_Trj_U/middle_data/individual(cleaned)", 
                        pattern = "\\.csv$", full.names = TRUE)

# 各CSVファイルを読み込む
for (file in csv_files) {
  # ファイル名（拡張子除く）を変数名とする
  var_name <- tools::file_path_sans_ext(basename(file))
  
  # assign関数を使って、変数名で指定した名前の変数にデータを代入する
  assign(var_name, read_csv(file), envir = .GlobalEnv)
}
```
##Outlier removal
#Outlier removal1: Remove records with mismatched years
```{r}
df_names <- paste0("GPS", 2013:2022)

for(df_name in df_names) {
  df <- get(df_name)
  df_year <- as.numeric(substr(df_name, 4, 7))
  df_clean <- df[year(df$time) == df_year, ]
  assign(df_name, df_clean, envir = .GlobalEnv)
}
```

#Outlier removal2: Sort data in chronological order
```{r}
for (year in 2013:2022) {
  df_name <- paste0("GPS", year)
  df <- get(df_name)
  df <- arrange(df, label, time)
  assign(df_name, df)
}
```

#Outlier removal3: Remove impossible latitude/longitude values
```{r}
year_list <- c("GPS2013", "GPS2014", "GPS2015", "GPS2016", "GPS2017", "GPS2018", "GPS2019", "GPS2020", "GPS2021", "GPS2022")
min_lon <- -180
max_lon <- 180
min_lat <- -90
max_lat <- 90

for (i in seq_along(year_list)) {
  df <- get(year_list[i]) 
  filtered_df <- df %>%
    filter(lon >= min_lon & lon <= max_lon,
           lat >= min_lat & lat <= max_lat)
  assign(year_list[i], filtered_df)
}
```


##Trip segmentation
#CalculateNestDistance
```{r}
# Location of Kabushima island
location_nest <- c(141.5576031, 40.5386062)

year_list <- c("GPS2013", "GPS2014", "GPS2015", "GPS2016", "GPS2017", "GPS2018", "GPS2019", "GPS2020", "GPS2021", "GPS2022")

for (year in year_list) {
  data <- get(year)
  data$Nest_Distance <- distGeo(location_nest, cbind(data$lon, data$lat))
  data$Nest_Distance <- data$Nest_Distance
  assign(year, data)
}
```


#Calculate speed and dt (time interval between consecutive data points)
```{r}
year_list <- c("GPS2013","GPS2014","GPS2015","GPS2016","GPS2017","GPS2018","GPS2019","GPS2020","GPS2021","GPS2022")

for (year in year_list) {
data_out_all <- data.frame()
data0 <- get(year)
id_list <- unique(data0$label)
n_id <- length(id_list)

for(i in 1:n_id) {
data <- data0%>%
  filter(label == id_list[i])
data_out <- data.frame()

n <- nrow(data)

Lon <- data$lon
Lat <- data$lat

location1 <- cbind(Lon[1:n-1], Lat[1:n-1])
location2 <- cbind(Lon[2:n], Lat[2:n])

Step_Length <- distGeo(location1, location2)
data$time <- as.POSIXct(data$time, format="%Y/%m/%d, %H:%M:%S")
dt <- data$time %>%
  diff() %>%
  as.numeric(unit=("secs"))

speed <- Step_Length/dt

Step_Length <- c(Step_Length, NA)
dt2 <- c(dt, NA)
speed2 <- c(speed, NA)*3.6
data_out <- mutate(data, speed2, dt2)
data_out_all <- bind_rows(data_out_all, data_out)
}
assign(paste0("speed", year), data_out_all)
}
```

#Speed range in each year
It include the unlikely value, which means it still including outlyer
```{r}
df_names <- paste0("speedGPS", 2013:2022)

for (df_name in df_names) {
  df <- get(df_name)
  df <- df[complete.cases(df$speed2) & !is.infinite(df$speed2), ]
  assign(df_name, df, envir = .GlobalEnv)
  print(paste0(df_name, " speed2 range: ", range(df$speed2, na.rm = TRUE)))
}
```

#Trip segmentation
```{r}
#Parameter
cut_distance <- 1000 #[m] Minimum distance of trip
cut_duration <- 60*120 #[sec] Minimum duration of trip
cut_duration %>% as.numeric(unit = "secs") 

year_list <- c("GPS2013", "GPS2014", "GPS2015", "GPS2016", "GPS2017", "GPS2018", "GPS2019", "GPS2020", "GPS2021", "GPS2022")

out_data <- tibble()

for (year in year_list) {
data0 <- get(year)  
out_data_year <- tibble()
id_list <- unique(data0$label)
n_id <- length(id_list)

for(i in 1:n_id){
data <- data0 %>%
  filter(label==id_list[i])
nest_out_index <- as.numeric(data$Nest_Distance > cut_distance)
trip_section0 <- nest_out_index %>%
    rle2(indices = TRUE) %>%
      as.data.frame() %>%
        filter(value == 1)

trip_duration <- (data[trip_section0$stop,]$time - data[trip_section0$start,]$time) %>% as.numeric(unit = "secs")
trip_section0 <- trip_section0 %>% mutate(trip_duration = trip_duration)

trip_section0 <- filter(trip_section0, trip_duration > cut_duration)

trip_ID <- rep(0, nrow(data))

if(nrow(trip_section0)>0){
for(j in 1:nrow(trip_section0)){
  trip_ID[trip_section0[j,]$start:trip_section0[j,]$stop] <- j
  
}
}

data$Trip_ID <- trip_ID
out_data <- bind_rows(out_data,data)

out_data <- mutate(out_data,ID = paste(out_data$year,out_data$foot,out_data$Trip_ID,"trip", sep="_"))
}
assign(year, out_data)
write_csv(out_data, paste0("~/Documents/yodalab/data/middle_data/Added/No1_trip_added/", year, ".csv"))
out_data <- tibble()
}
```

#Remove the data with speed > 100km/h
Likely the noize
```{r}
# Input parameters
MAX_SPEED <-100 #[km/hour]


lat_max <-50
lat_min <-20
lon_min <-100
lon_max <-180

input_dir <- "~/Documents/yodalab/data/middle_data/Added/No1_trip_added/"

output_dir <- "~/Documents/yodalab/data/middle_data/Added/No1_trip_added/SDL_filtered/"

years <- 2013:2022

for (year in years) {
  in_file <- paste0(input_dir, "GPS", year, ".csv")
  out_file <- paste0(output_dir, "GPS", year, ".csv")
  data_all <- c()
  data0 <- read_csv(in_file)
  no_col_of_raw_data <- ncol(data0)
  data0 <- data0 %>% 
             filter(lon < lon_max) %>%
             filter(lon > lon_min) %>%
             filter(lat < lat_max) %>%
             filter(lat > lat_min) 
  data0 <- mutate(data0, id = paste(data0$label,data0$year))
  id_list <- unique(data0$id)
  n_id <- length(id_list)

  # j: individual ID
  for(j in 1:n_id){
    data <- filter(data0, id == id_list[j])
    data <- data[order(data$time),]
    data$qi <- rep(10,  nrow(data))
    names(data)[which(names(data)=="time")] <- "DateTime"
    vmax <- MAX_SPEED
    data <- ddfilter(as.data.frame(data), vmax=vmax, qi=4, ia=90, method=2)
    n_data <- nrow(data)
    names(data)[which(names(data)=="DateTime")] <- "time"
    data <- data[,1:no_col_of_raw_data]
    data_all <- rbind(data_all,data)
  }
  write_csv(data_all, out_file)
}
```

#Calculate speed and dt (time interval between consecutive data points) again
```{r}
rm(list=ls())
csv_files <- list.files(path = "~/Documents/yodalab/data/middle_data/Added/No1_trip_added/SDL_filtered/", 
                        pattern = "\\.csv$", full.names = TRUE)

for (file in csv_files) {
  var_name <- tools::file_path_sans_ext(basename(file))
  assign(var_name, read_csv(file), envir = .GlobalEnv)
}
year_list <- c("GPS2013","GPS2014","GPS2015","GPS2016","GPS2017","GPS2018","GPS2019","GPS2020","GPS2021","GPS2022")

for (year in year_list) {
data_out_all <- data.frame()
data0 <- get(year)
id_list <- unique(data0$label)
n_id <- length(id_list)

for(i in 1:n_id) {
data <- data0%>%
  filter(label == id_list[i])
data_out <- data.frame()
n <- nrow(data)
Lon <- data$lon
Lat <- data$lat
location1 <- cbind(Lon[1:n-1], Lat[1:n-1])
location2 <- cbind(Lon[2:n], Lat[2:n])
Step_Length <- distGeo(location1, location2)
data$time <- as.POSIXct(data$time, format="%Y/%m/%d, %H:%M:%S")
dt <- data$time %>%
  diff() %>%
  as.numeric(unit=("secs"))
speed <- Step_Length/dt
Step_Length <- c(Step_Length, NA)
dt2 <- c(dt, NA)
speed2 <- c(speed, NA)*3.6
data_out <- mutate(data, speed2, dt2)
data_out_all <- bind_rows(data_out_all, data_out)
}
assign(paste0("speed", year), data_out_all)
}
```

#Speed range in each year2
```{r}
df_names <- paste0("speedGPS", 2013:2022)
for (df_name in df_names) {
  df <- get(df_name)
  df <- df[complete.cases(df$speed2) & !is.infinite(df$speed2), ]
  assign(df_name, df, envir = .GlobalEnv)
  print(paste0(df_name, " speed range: ", range(df$speed2, na.rm = TRUE)))
}
```

#Remove the non-trip GPS data
+ Removal of trips containing a time interval between consecutive data points of 1800 seconds (30 minutes) or more
```{r}
year_list <- c("speedGPS2013", "speedGPS2014", "speedGPS2015", "speedGPS2016", "speedGPS2017", "speedGPS2018", "speedGPS2019", "speedGPS2020", "speedGPS2021", "speedGPS2022")

for (year in year_list) {
data <- get(year)

data <- data %>%
  dplyr::filter(!grepl("0_trip$", ID))

ids_to_remove <- data %>%
  filter(dt2 >= 1800) %>%
  pull(ID)

data <- data %>%
  filter(!ID %in% ids_to_remove)
assign(paste0("edit", year), data)

showid <-  unique(ids_to_remove)
  # "trip0"で終わるIDは除外
showid <- showid[!grepl("0_trip$", showid)]
# 削除されたIDを表示
  print(paste(year, "removed IDs:", paste(showid, collapse = ", ")))
}
```

#Remove trips where maximum distance from nest is less than 20 km
```{r}
year_list <- c("editspeedGPS2013", "editspeedGPS2014", "editspeedGPS2015", "editspeedGPS2016", "editspeedGPS2017", "editspeedGPS2018", "editspeedGPS2019", "editspeedGPS2020", "editspeedGPS2021", "editspeedGPS2022")


for (year in year_list) {
  data0 <- get(year)
  id_list <- unique(data0$ID)
  out_data <- tibble()
  removed_ids <- c()

  for(i in unique(id_list)) {
    data <- filter(data0, ID == i)
    
    # If the max Nest_Distance is greater than 20000, keep the data. Else, store the ID and move to the next loop iteration
    if(max(data$Nest_Distance, na.rm = TRUE) <= 20000) {
      removed_ids <- c(removed_ids, i)
      next
    }
    out_data <- bind_rows(out_data, data)
  }

  if (length(removed_ids) > 0) {
    print(paste(year, "20000m removed IDs:", paste(removed_ids, collapse = ", ")))
  }
  year_num <- str_extract(year, "\\d+")
  assign(paste0("longtrip", year_num), out_data)
  out_place <- paste0("~/Documents/yodalab/data/middle_data/Added/No2_wrongtrip_removed/longtrip", year_num, ".csv")
  write_csv(out_data, out_place)
}
```

##Cleaning
#Read again
```{r}
rm(list=ls())
csv_files <- list.files(path = "~/Documents/yodalab/data/middle_data/Added/No2_wrongtrip_removed/", 
                        pattern = "\\.csv$", full.names = TRUE)
for (file in csv_files) {
  var_name <- tools::file_path_sans_ext(basename(file))
  assign(var_name, read_csv(file), envir = .GlobalEnv)
}
```



#Outlier removal4: Retain only data from April to July, and remove any trips that contain data points outside this period
```{r}
range(longtrip2019$time)
range(longtrip2021$time)

year_list <- c("longtrip2013", "longtrip2014", "longtrip2015", "longtrip2016", "longtrip2017", "longtrip2018", "longtrip2019", "longtrip2020", "longtrip2021", "longtrip2022")

for (year in year_list) {
  data <- get(year)
  data$time <- as.POSIXct(data$time, format = "%Y-%m-%d %H:%M:%S")
  
  ids_to_remove <- data %>%
    filter(!(month(time) >= 5 & month(time) <= 7)) %>%
    pull(ID) %>%
    unique()

  data <- data %>%
    filter(!ID %in% ids_to_remove)
  
  print(ids_to_remove)
  assign(year, data)
}

#除かれたもの
#[1] "2019_09214934_2_trip" "2019_09214934_3_trip"
#[3] "2019_9A90224_1_trip"  "2019_9A90224_2_trip" 
#[5] "2019_9A90224_3_trip"  "2019_9A94429_2_trip" 
#[7] "2019_9A94429_3_trip"  "2019_9B16224_1_trip" 
#[9] "2019_9B16224_2_trip" 
#[1] "2021_9A94429_1_trip"
```

#Outlyer removal5: Remove trips where the last data point is more than 3 km from the nest
```{r}
year_list <- c("longtrip2013","longtrip2014","longtrip2015","longtrip2016","longtrip2017","longtrip2018","longtrip2019","longtrip2020","longtrip2021","longtrip2022")

for (year in year_list) {
  data <- get(year)
  
  last_distance_per_id <- data %>% 
    group_by(ID) %>% 
    summarise(last_distance = last(Nest_Distance))
  
  ids_to_remove <- last_distance_per_id %>% 
    filter(last_distance > 3000) %>% 
    pull(ID)
  
  rows_to_remove <- data %>% 
    filter(ID %in% ids_to_remove)
  
  print(paste("Number of rows removed for", year, ":", nrow(rows_to_remove)))
  
  data_cleaned <- data %>% 
    filter(!(ID %in% ids_to_remove))
  
  assign(year, data_cleaned, envir = .GlobalEnv)
}

```

#Check individual number
```{r}
df_names <- c("longtrip2013", "longtrip2014", "longtrip2015", "longtrip2016", "longtrip2017", "longtrip2018", "longtrip2019", "longtrip2020", "longtrip2021", "longtrip2022")
unique_counts <- lapply(df_names, function(df_name) {
  df <- get(df_name)
  length(unique(df$foot))
})
result <- data.frame(
  DataFrame = df_names,
  UniqueIDs = unlist(unique_counts)
)
print(result)
```

##Visualize each individual
```{r}
out_place <- "~/Documents/yodalab/data/middle_data/Added/No2_wrongtrip_removed/Trip_plot/"
year_list <- c("longtrip2013", "longtrip2014", "longtrip2015", "longtrip2016", "longtrip2017", "longtrip2018", "longtrip2019", "longtrip2020", "longtrip2021", "longtrip2022")
#グラデーションできるように、palの中に色を10色入れる
pal <- c("#3B9AB2", "#56A6BA", "#71B3C2", "#9EBE91", "#D1C74C",
         "#E8C520", "#E4B80E", "#E29E00", "#EA5C00", "#F21A00")

for (year in year_list) {
  final_data <- get(year)
  id_list <- unique(final_data$ID)
  n_id <- length(id_list)
  for(i in 1:n_id){
    data1 <- final_data %>%
      filter(ID==id_list[i])
    
    fig<- ggplot(data1, aes(x=lon, y=lat, colour = Nest_Distance))+geom_point(size=0.5)+ scale_color_gradientn(colors = pal)+theme_light()

    ggsave(file = paste(out_place, id_list[i],".png", sep=""), plot=fig)
  }
}
```

##Output
```{r}
output_list <- c("longtrip2013","longtrip2014","longtrip2015","longtrip2016","longtrip2017","longtrip2018","longtrip2019","longtrip2020","longtrip2021","longtrip2022")

out_place <- "~/Documents/yodalab/data/middle_data/Added/No2_wrongtrip_removed/"

for (year in output_list) {
  output_data <- get(year)

#  output_data <- output_data %>% 
#  dplyr::select(time, lon, lat, id, label, age, sex, year, logger, type, Nest_Distance, Trip_ID, ID, speed2, dt2)

  write_csv(output_data, paste0(out_place, year, ".csv"))
}
```