---
title: "Code8_Coast_Deviation"
output: html_document
date: "2026-09-01"
---
---
title: "New12_目的地を巣以外にする"
output: html_document
date: "2026-07-14"
---

#library
```{r}
rm(list=ls())
library(sf)
library(rnaturalearth)
library(rnaturalearthdata)
library(ggplot2)
library(readr)
library(dplyr)
library(sf)
library(ggspatial)
library(circular)
library(ncdf4)
library(maps)
library(lubridate)
library(viridis)
library(purrr)
library(extrafont)
library(patchwork)
library(tidyr)
library(tidyverse)
library(gridExtra)
library(lme4)
library(ggeffects)
library(geosphere)
library(here)
library(ggplot2)
library(RColorBrewer)
library(slider)
library(emmeans)
```

#data
```{r}
rm(list=ls())

#Read final data after all claning
all_sea_data <- read_csv("~/Documents/yodalab/data/final_data/Original_from_yodalab/Windcompensation/all_sea_data_wind_recalculated.csv")

all_sea_data <- all_sea_data[!is.na(all_sea_data$age), ]

#Read date before HMM classification (before removing coast)
out_dir <- "~/Documents/yodalab/data/middle_data/Added/No3_5min_resampled/fixed/"
all_data <- map_dfr(
  file.path(out_dir, paste0("GPS", 2013:2022, ".csv")),
  read_csv
)
```


#Recover the coastal flight
Coastal flights have already been excluded from the post-HMM data, so the current dataset contains only the homing portion. Here, we want to add back the GPS data from Code 3 before coastal flights were excluded.
We can do this by adding all GPS points that follow the homing portion.
```{r}
# Find the last timestamp for each ID (trip)
last_time_by_ID <- all_sea_data %>%
  group_by(ID) %>%
  summarise(
    last_time = max(time, na.rm = TRUE),
    .groups = "drop"
  )

# Find the continuation of each trip in all_data
# Keep only IDs that are present in all_sea_data
continuation_data <- all_data %>%
  inner_join(last_time_by_ID, by = "ID")

# Keep only points recorded after the last timestamp in all_sea_data
continuation_data <- continuation_data %>%
  filter(time > last_time)

# Remove the temporary last_time column
continuation_data <- continuation_data %>%
  dplyr::select(-last_time)


# Label the added points as "coast" for HomingID, State, and original_state
continuation_data <- continuation_data %>%
  mutate(HomingID = "coast") %>% 
  mutate(State = "coast") %>% 
  mutate(original_state = "coast")


# Add the continuation data back to all_sea_data
all_sea_data_extended <- bind_rows(all_sea_data, continuation_data)
all_sea_data_extended <- all_sea_data_extended %>% arrange(ID, time)
```

#Recalculate distance to the coast
```{r}
# Convert the coastline into a dense set of points
japan_map <- map_data("world", region = "Japan")
japan_map_interp <- japan_map %>%
  group_by(group) %>%
  do(data.frame(
    long = approx(seq_along(.$long), .$long, n = 200 * length(.$long))$y,
    lat = approx(seq_along(.$lat), .$lat, n = 200 * length(.$lat))$y
  )) %>%
  dplyr::filter(long >= 139, long <= 145.8260, lat >= 37, lat <= 44) %>%
  ungroup()

# Define a function that returns the nearest coastline point
# and its distance for each GPS location
nearest_point_distance <- function(lon, lat) {

  distances <- distHaversine(
    matrix(c(lon, lat), ncol = 2),
    matrix(c(japan_map_interp$long, japan_map_interp$lat), ncol = 2)
  )

  min_index <- which.min(distances)

  tibble(
    coast_lon = japan_map_interp$long[min_index],
    coast_lat = japan_map_interp$lat[min_index],
    coast_distance = distances[min_index]
  )
}

# Remove existing coast-related columns before recalculating them
all_sea_data_extended <- all_sea_data_extended %>%
  dplyr::select(-any_of(c("coast_lon", "coast_lat", "coast_distance")))

nearest <- pmap_dfr(
  all_sea_data_extended[, c("lon", "lat")],
  nearest_point_distance
)

all_sea_data_extended <- bind_cols(
  all_sea_data_extended,
  nearest
)
```

#Save
```{r}
write.csv(all_sea_data_extended, "~/Documents/yodalab/data/middle_data/Added/No6_coast_readded_byreview/all_data_coast_distance.csv")
```

#Read
```{r}
all_sea_data_extended <- read_csv("~/Documents/yodalab/data/middle_data/Added/No6_coast_readded_byreview/all_data_coast_distance.csv")
```

# Identify the first point where the bird transitions to coastal movement
```{r}
# Keep all columns up to coast_distance and remove columns to the right
all_sea_data_extended <- all_sea_data_extended %>%
  dplyr::select(1:coast_distance)

# 1. Calculate approach speed toward the coast for each ID
all_sea_data_extended <- all_sea_data_extended %>%
  arrange(ID, time) %>%
  group_by(ID) %>%
  mutate(
    dt_sec = as.numeric(
      difftime(time, lag(time), units = "secs")
    ),
    # Positive values = approaching the coast
    # Values near 0 = little change in distance to the coast
    # Negative values = moving away from the coast
    coast_approach_speed =
      (lag(coast_distance) - coast_distance) / dt_sec
  ) %>%
  ungroup()


# 2. Calculate approach speed before and after each potential transition point
all_sea_data_extended <- all_sea_data_extended %>%
  arrange(ID, time) %>%
  group_by(ID) %>%
  mutate(
    
    # Mean approach speed over the previous three points
    approach_speed_before =
      (
        lag(coast_approach_speed, 1) +
        lag(coast_approach_speed, 2) +
        lag(coast_approach_speed, 3)
      ) / 3,
    
    # Mean approach speed over the current and next two points
    approach_speed_after =
      (
        coast_approach_speed +
        lead(coast_approach_speed, 1) +
        lead(coast_approach_speed, 2)
      ) / 3,
    
    # Amount by which approach speed toward the coast decreases
    approach_speed_drop =
      approach_speed_before - approach_speed_after
  ) %>%
  ungroup()


# 3. Standard criterion:
#    For each ID, identify the first point within 2 km of the coast
#    where the approach speed toward the coast decreases
closest_coast_point <- all_sea_data_extended %>%
  filter(
    !is.na(approach_speed_drop),
    coast_distance < 2000,        # Within 2 km of the coast
    approach_speed_before > 0,    # Previously approaching the coast
    approach_speed_drop > 0       # Approach speed subsequently decreased
  ) %>%
  arrange(ID, time) %>%
  group_by(ID) %>%
  slice(1) %>%
  ungroup() %>%
  dplyr::select(
    ID,
    closest_coast_time = time,
    closest_coast_lon = lon,
    closest_coast_lat = lat,
    coast_distance_at_transition = coast_distance
  )


# 4. Fallback:
#    For IDs where no closest-coast point was identified using the standard criterion,
#    use the final point as the closest-coast point if it is:
#    ・more than 5 km from the nest
#    ・within 2 km of the coast
last_point_fallback <- all_sea_data_extended %>%
  arrange(ID, time) %>%
  group_by(ID) %>%
  slice_tail(n = 1) %>%
  ungroup() %>%
  filter(
    Nest_Distance > 5000,        # More than 5 km from the nest
    coast_distance < 2000        # Within 2 km of the coast
  ) %>%
  anti_join(
    closest_coast_point %>% dplyr::select(ID),
    by = "ID"
  ) %>%
  dplyr::select(
    ID,
    closest_coast_time = time,
    closest_coast_lon = lon,
    closest_coast_lat = lat,
    coast_distance_at_transition = coast_distance
  )


# 5. Combine points identified by the standard criterion and fallback
closest_coast_point <- bind_rows(
  closest_coast_point,
  last_point_fallback
)


# 6. Add transition-point information to all_sea_data_extended
all_sea_data_extended <- all_sea_data_extended %>%
  left_join(
    closest_coast_point,
    by = "ID"
  )
```

#Use the nest as the closest-coast goal when no separate coastal transition is identified
```{r}
# 1. Nest location

location_nest_data <- data.frame(
  lon = 141.5576031,
  lat = 40.5386062
)

nest_lon <- location_nest_data$lon[1]
nest_lat <- location_nest_data$lat[1]

# 2. Set the closest-coast point as the goal
#    If a coastal transition point is identified: goal = closest coast
#    If no coastal transition point is identified: goal = nest

all_sea_data_extended <- all_sea_data_extended %>%
  mutate(
    goal_lon = if_else(
      !is.na(closest_coast_lon),
      closest_coast_lon,
      nest_lon
    ),
    goal_lat = if_else(
      !is.na(closest_coast_lat),
      closest_coast_lat,
      nest_lat
    ),
    goal = if_else(
      !is.na(closest_coast_lon),
      "coast", "nest"
    )
  )

# 7. Calculate goal deviation for each ID

goal_summary <- all_sea_data_extended %>%
  distinct(ID, age, label, foot, goal, goal_lon, goal_lat) %>%
  rowwise() %>%
  mutate(
    goal_deviation_straight = distHaversine(
      c(goal_lon, goal_lat),
      c(nest_lon, nest_lat)
    )
  ) %>%
  ungroup()
```